### 2016-09-15 Chapters/MathematicalResults.tex
Polish

@@

\begin{defin}[Coercivity]\label{def:coercivity}
  A bilinear form on a normed vector space $V$ is \emph{coercive},
  or \emph{elliptic}, if there is a constant $\alpha >0$ such that
  $\forall u \in V$:
  \[
  a(u, u) \geq \alpha\lVert u\rVert^2
  \]
\end{defin}
Coercivity implies that no eigenvalue of the linear operator associated to the
bilinear form can be zero, hence its invertibility.~\autocite{Ern2004-oo}

@@

\section{Derivation of the IEF equation}
I will show a detailed derivation of the \acrshort{IEF} equation
for the \acrshort{PCM} transmission problem Eq.~\eqref{eq:transmission}:
\begin{subequations}
  \begin{align}[left={\empheqlbrace}]
  \Li u &= \nabla^2 u = -4\pi\rhoi \,\, \text{in}\,\, \Omegai \\
  \Le u &= 0 \,\, \text{in}\,\, \Omegae \\
  [u] &= \ue - \ui = 0 \,\, \text{on}\,\, \Gamma \\
[\partial_L u] &= \partiale u - \partiali u = 0 \,\, \text{on}\,\, \Gamma \\
|u(\vect{r})| &\leq C \|\vect{r} \|^{-1} \,\,\text{for}\,\,\| \vect{r} \|\rightarrow\infty
\end{align}
\end{subequations}

We first state two important results in the theory of boundary integral
equations.
\begin{lemma}[Properties of the boundary integral operators]
  The integral operators introduced in Eqs.~\eqref{eq:S}-\eqref{eq:D-dagger}
  enjoy the following properties~\autocite{Hsiao2008-xb, Sauter2011-an}:
  \begin{enumerate}
      \item on $L^2(\Gamma)$, $\bi{S}_\star$ is self-adjoint,
        $\bi{D}^\dagger_\star$ is the adjoint operator of
        $\bi{D}_\star$.
      \item The commutation relations hold:
        \begin{alignat}{2}
          \bi{D}_\star\bi{S}_\star = \bi{S}_\star\bi{D}^\dagger_\star, \quad&
          \bi{S}_\star\bi{D}_\star = \bi{D}^\dagger_\star\bi{S}_\star
        \end{alignat}
      \item The boundary integral operators are continuous
        mappings between Sobolev spaces of fractional order:
        \begin{subequations}
          \begin{align}
   \bi{S}_\star &: H^{-\frac{1}{2}}(\Gamma) \rightarrow H^{\frac{1}{2}}(\Gamma) \\
   \bi{D}_\star &: H^{\frac{1}{2}}(\Gamma) \rightarrow H^{\frac{1}{2}}(\Gamma) \\
   \bi{D}^\dagger_\star &: H^{-\frac{1}{2}}(\Gamma) \rightarrow H^{-\frac{1}{2}}(\Gamma)
          \end{align}
        \end{subequations}
      \item The operator $\bi{S}_\star$ is coervice and admits a
        continuous inverse in the aforementioned Sobolev spaces.
      \item The operators $\lambda - \bi{D}_\star$ and $\lambda -
        \bi{D}^\dagger_\star$ with $\lambda \in (-2\pi, +\infty)$
        admit a continuous inverse in the aforementioned
        Sobolev spaces.
  \end{enumerate}
\end{lemma}

@@

To derive an equation for $\sigma$, we set up a system of equations containing
the jump conditions and the integral representations of the reaction and auxiliary potentials:
\begin{subequations}
\begin{empheq}[left={\empheqlbrace}]{align}
    &\bi{S}_\mathrm{i}(\partiali \xi) - \bi{D}_\mathrm{i}(\xi_\mathrm{i})
    = \frac{1}{2}\xi_\mathrm{i} \label{eq:first} \\
    &\bi{S}_\mathrm{e}(\partiale \xi) - \bi{D}_\mathrm{e}(\xi_\mathrm{e})
    = -\frac{1}{2}\xi_\mathrm{e} \label{eq:second} \\
  &\bi{S}_\mathrm{e}(\partiale h) - \bi{D}_\mathrm{e}(h_\mathrm{e})
  = -\frac{1}{2}h_\mathrm{e} \label{eq:third} \\
  &\xi_\mathrm{i} - \xi_\mathrm{e} = h_\mathrm{e} - h_\mathrm{i}
  \label{eq:fourth} \\
  &\partiali \xi - \partiale \xi = \partiale h - \partiali h
  \label{eq:fifth}
\end{empheq}
\end{subequations}
The final ingredient is the \gls{DtN} map, which can be derived by employing Eq.~\eqref{eq:stat3}
to the Newton potential:
\begin{equation}
  \phi(\vect{r}) = (\bi{N}\rhoi)(\vect{r}) = \int_{\mathbb{R}^3}\Gi\rhoi(\vect{r}^\prime)\diff \vect{r}^\prime
   = \left.h(\vect{r})\right|_{\Omegai}
\end{equation}
which is equal, in $\Omegai$, to the auxiliary potential $h(\vect{r})$.
The \acrshort{DtN} map is thus given as:
\begin{equation}\label{eq:regular-DtN}
   \left(\frac{1}{2} - \bi{D}_\mathrm{i}\right)\phi_\mathrm{i}
      +\bi{S}_\mathrm{i}(\partiali \phi) = 0
\end{equation}
With this last ingredient at hand, algebraic manipulations lead to the \acrshort{IEF}
equation:
\begin{equation}
  \left[ \bi{S}_\mathrm{e}\left(2\pi + \bi{D}^\dagger_\mathrm{i}\right)
  +
  \left(2\pi - \bi{D}_\mathrm{e}\right)\bi{S}_\mathrm{i}
  \right]\sigma =
  -\left[\left(2\pi-\bi{D}_\mathrm{e}\right)
  -\bi{S}_\mathrm{e}\bi{S}_\mathrm{i}^{-1}
  \left(2\pi-\bi{D}_\mathrm{i}\right)
  \right]\esp,
  \tag{\ref{eq:full-IEF} from Chapter \ref{ch:CSM}}
\end{equation}

@@

In this Section, I will show how one can reformulate the transmission problem
in a variational fashion.
Such a formulation allows for a larger vector space, with weaker regularity
conditions, to be explored as solution space for the problem.
We will follow the exposition of \citeauthor{Ern2004-oo} quite closely
in introducing the \emph{weak} formulation of \glspl{PDE}.
For simplicity, we assume Dirichlet boundary conditions for a conductor,
\ie the basic assumption behind \acrshort{COSMO}.
The \emph{strong} formulation of the electrostatic problem then reads:
\begin{equation}\label{eq:poisson-cosmo}
  \nabla^2\esp = -4\pi\rho,
  \quad \esp \in \mathcal{C}^2_0(\Omegai),
\end{equation}
where $\mathcal{C}^2_0(\Omegai)$ is the vector space of twice
continuously differentiable functions in $\Omegai$ with null trace on
$\Gamma$.
We can relax this regularity requirement on $\Psi$ by introducing the
\emph{Hilbert Sobolev space} of test functions $H_0^1(\Omegai)$:
\begin{equation}
  H_0^1(\Omegai) = \lbrace
  f : \Omegai \rightarrow \mathbb{R}
  | f, \nabla f \in L^2(\Omegai), f_\Gamma = 0
  \rbrace.
\end{equation}
Projecting the differential problem onto this space and using the
$\eta\nabla^2\gamma = \nabla\cdot(\eta\nabla\gamma) - \nabla\eta\cdot\nabla\gamma$ identity
one obtains the \emph{weak} formulation of the differential problem:
\begin{equation}\label{eq:poisson-weak}
  \left\{
  \begin{aligned}
    &\text{Seek $\Psi \in H_0^1(\Omegai)$ such that:}\\
    &(\nabla\zeta, \nabla\Psi) =
    -4\pi(\zeta, \rho) \quad
    \forall \zeta\in H_0^1(\Omegai)
  \end{aligned}
  \right.
\end{equation}
The form $(\nabla\cdot, \nabla\cdot)$ is \emph{bilinear} and continuous
in $H_0^1(\Omegai)\times H_0^1(\Omegai)$, while $-4\pi(\cdot, \rho)$ is
\emph{linear} and continuous in $H_0^1(\Omegai)$.
We rewrite Eq.~\eqref{eq:poisson-weak} in the abstract form:
\begin{equation}
  \left\{
  \begin{aligned}
    &\text{Seek $u \in V$ such that:}\\
    &\forall v \in V \,\,\,\,\,
    a(u, v) = b(v)
  \end{aligned}
  \right.
  \tag{\ref{eq:weak}}
\end{equation}
and state the following fundamental results:
\begin{lemma}[Lax--Milgram]
  If the bilinear form $a$ is \emph{continous} and \emph{coercive}
  in $V$, then, for any continuous linear form $b$, Problem~\eqref{eq:weak} is
  \emph{well-posed}.
\end{lemma}
\begin{corollary}[Variational property]
  If the bilinear form is \emph{symmetric} and \emph{positive}
  the unique solution to Problem \eqref{eq:weak} is
  the \emph{unique minimum} on $V$ of the functional:
  \[
  \mathcal{F}(u) = \frac{1}{2}a(u, u) - b(u)
  \]
\end{corollary}

@@

\todo[inline]{Write this section}
Approximability, conformity, consistency and asymptotic consistency
of the approximation setting, $\inf$-$\sup$ condition, Céa's Lemma
and generalized $\inf$-$\sup$ condition.~\autocite{Ern2004-oo}

@@

\begin{equation}
  \begin{pmatrix}
   \sigma^\mathrm{tot} \\
   \kappa^\mathrm{tot}
  \end{pmatrix}
  =
  \begin{pmatrix}
   \sigma_\mathrm{HF} \\
   \kappa_\mathrm{HF}
  \end{pmatrix}
  +
  \begin{pmatrix}
   \sigma \\
   \kappa
  \end{pmatrix}
\end{equation}
and similarly for the sources:
\begin{equation}
\begin{aligned}
  \begin{pmatrix}
   \esp(\tamp{}, \tbar{}) \\
   \zeta(\tamp{}, \tbar{})
  \end{pmatrix}
  &=
  \begin{pmatrix}
   \esp_\mathrm{HF} \\
   \zeta_\mathrm{HF}
  \end{pmatrix}
  +
  \begin{pmatrix}
    \esp_{\mathrm{N}}(\tamp{}, \tbar{}) \\
    \zeta_{\mathrm{N}}(\tamp{}, \tbar{})
  \end{pmatrix} \\
  &=
  \begin{pmatrix}
  \braket{\mathrm{HF} | \esp  | \mathrm{HF}} \\
  \braket{\mathrm{HF} | \zeta | \mathrm{HF}}
  \end{pmatrix}
  +
  \begin{pmatrix}
   \braket{\mathrm{HF} | \tilde{\esp} | \mathrm{HF}}
  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{\esp} | \mathrm{HF}} \\
   \braket{\mathrm{HF} | \tilde{\zeta} | \mathrm{HF}}
  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{\zeta} | \mathrm{HF}}
  \end{pmatrix}
\end{aligned}
\end{equation}
The \acrshort{BCH} expansions of the source terms have been
rewritten as:
\begin{subequations}
  \begin{align}
  \overline{\esp} &= \esp + \tilde{\esp} = \esp + \commutator{\esp}{T}
  + \frac{1}{2}\commutator{\commutator{\esp}{T}}{T} \\
  \overline{\kappa} &= \kappa + \tilde{\kappa} = \kappa + \commutator{\kappa}{T}
  + \frac{1}{2}\commutator{\commutator{\kappa}{T}}{T}
  \end{align}
\end{subequations}
taking advantage of the fact that both $\esp$ and $\kappa$ are
nondiagonal, one-electron operators and hence their commutator
expansions truncate at the third term. Moreover, normal-ordering of the
operators has been introduced.~\autocite{Crawford2000-ey, Shavitt2009-mr}
This is equivalent to applying a shift to the correlation part of the
source term one-electron operators to remove the reference source terms:
\begin{equation}
  \begin{aligned}
  \begin{pmatrix}
    \esp_{\mathrm{N}}(\tamp{}, \tbar{}) \\
    \zeta_{\mathrm{N}}(\tamp{}, \tbar{})
  \end{pmatrix}
  &=
  \begin{pmatrix}
   \braket{\mathrm{HF} | \tilde{\esp} | \mathrm{HF}}
  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{\esp} | \mathrm{HF}} \\
   \braket{\mathrm{HF} | \tilde{\zeta} | \mathrm{HF}}
  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{\zeta} | \mathrm{HF}}
  \end{pmatrix}
  \\
  &=
  \begin{pmatrix}
    \braket{\mathrm{HF} | \overline{\esp} - \esp_\mathrm{HF} | \mathrm{HF}}
  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{\esp} - \esp_\mathrm{HF} | \mathrm{HF}} \\
  \braket{\mathrm{HF} | \overline{\zeta} - \zeta_\mathrm{HF} | \mathrm{HF}}
  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{\zeta} -
  \zeta_\mathrm{HF}| \mathrm{HF}}
  \end{pmatrix}
\end{aligned}
\end{equation}

>>>>

\begin{epigraphs}
\qitem{\textitalian{Non temerai i terrori della notte non temerai il terrore}}{--- \textsc{CCCP}}
\end{epigraphs}

@@

\begin{defin}[Coercivity]\label{def:coercivity}
  A bilinear form on a normed vector space $V$ is \emph{coercive},
  or \emph{elliptic}, if there is a constant $\alpha >0$ such that
  $\forall u \in V$:
  \[
  a(u, u) \geq \alpha\lVert u\rVert^2
  \]
\end{defin}
Coercivity implies that no eigenvalue of the linear operator associated to the
bilinear form can be zero, hence its invertibility.\autocite{Ern2004-oo}

@@

\section{Derivation of the IEF equation}
I will show a detailed derivation of the \acrshort{IEF} equation
for the \acrshort{PCM} transmission problem Eq. \eqref{eq:transmission}:
\begin{subequations}
  \begin{align}[left={\empheqlbrace}]
  \Li u &= \nabla^2 u = -4\pi\rhoi \,\, \text{in}\,\, \Omegai \\
  \Le u &= 0 \,\, \text{in}\,\, \Omegae \\
  [u] &= \ue - \ui = 0 \,\, \text{on}\,\, \Gamma \\
[\partial_L u] &= \partiale u - \partiali u = 0 \,\, \text{on}\,\, \Gamma \\
|u(\vect{r})| &\leq C \|\vect{r} \|^{-1} \,\,\text{for}\,\,\| \vect{r} \|\rightarrow\infty
\end{align}
\end{subequations}

We first state two important results in the theory of boundary integral
equations.
\begin{lemma}[Properties of the boundary integral operators]
  The integral operators introduced in Eqs. \eqref{eq:S}--\eqref{eq:D-dagger}
  enjoy the following properties\autocite{Hsiao2008-xb, Sauter2011-an}:
  \begin{enumerate}
      \item on $L^2(\Gamma)$, $\bi{S}_\star$ is self-adjoint,
        $\bi{D}^\dagger_\star$ is the adjoint operator of
        $\bi{D}_\star$.
      \item The commutation relations hold:
        \begin{alignat}{2}
          \bi{D}_\star\bi{S}_\star = \bi{S}_\star\bi{D}^\dagger_\star, \quad&
          \bi{S}_\star\bi{D}_\star = \bi{D}^\dagger_\star\bi{S}_\star
        \end{alignat}
      \item The boundary integral operators are continuous
        mappings between Sobolev spaces of fractional order:
        \begin{subequations}
          \begin{align}
   \bi{S}_\star &: H^{-\frac{1}{2}}(\Gamma) \rightarrow H^{\frac{1}{2}}(\Gamma) \\
   \bi{D}_\star &: H^{\frac{1}{2}}(\Gamma) \rightarrow H^{\frac{1}{2}}(\Gamma) \\
   \bi{D}^\dagger_\star &: H^{-\frac{1}{2}}(\Gamma) \rightarrow H^{-\frac{1}{2}}(\Gamma)
          \end{align}
        \end{subequations}
      \item The operator $\bi{S}_\star$ is coervice and admits a
        continuous inverse in the aforementioned Sobolev spaces.
      \item The operators $\lambda - \bi{D}_\star$ and $\lambda -
        \bi{D}^\dagger_\star$ with $\lambda \in (-2\pi, +\infty)$
        admit a continuous inverse in the aforementioned
        Sobolev spaces.
  \end{enumerate}
\end{lemma}

@@

To derive an equation for $\sigma$, we set up a system of equations containing
the jump conditions and the integral representations of the reaction and auxiliary potentials:
\begin{subequations}
\begin{empheq}[left={\empheqlbrace}]{align}
    &\bi{S}_\mathrm{i}(\partiali \xi) - \bi{D}_\mathrm{i}(\xi_\mathrm{i})
    = \frac{1}{2}\xi_\mathrm{i} \label{eq:first} \\
    &\bi{S}_\mathrm{e}(\partiale \xi) - \bi{D}_\mathrm{e}(\xi_\mathrm{e})
    = -\frac{1}{2}\xi_\mathrm{e} \label{eq:second} \\
  &\bi{S}_\mathrm{e}(\partiale h) - \bi{D}_\mathrm{e}(h_\mathrm{e})
  = -\frac{1}{2}h_\mathrm{e} \label{eq:third} \\
  &\xi_\mathrm{i} - \xi_\mathrm{e} = h_\mathrm{e} - h_\mathrm{i}
  \label{eq:fourth} \\
  &\partiali \xi - \partiale \xi = \partiale h - \partiali h
  \label{eq:fifth}
\end{empheq}
\end{subequations}
The final ingredient is the \gls{DtN} map, which can be derived by employing Eq. \eqref{eq:stat3}
to the Newton potential:
\begin{equation}
  \phi(\vect{r}) = (\bi{N}\rhoi)(\vect{r}) = \int_{\mathbb{R}^3}\Gi\rhoi(\vect{r}^\prime)\diff \vect{r}^\prime
   = \left.h(\vect{r})\right|_{\Omegai}
\end{equation}
which is equal, in $\Omegai$, to the auxiliary potential $h(\vect{r})$.
The \acrshort{DtN} map is thus given as:
\begin{equation}\label{eq:regular-DtN}
   \left(\frac{1}{2} - \bi{D}_\mathrm{i}\right)\phi_\mathrm{i}
      +\bi{S}_\mathrm{i}(\partiali \phi) = 0
\end{equation}
With this last ingredient at hand, algebraic manipulations lead to the \acrshort{IEF}
equation:
\begin{equation}
  \left[ \bi{S}_\mathrm{e}\left(2\pi + \bi{D}^\dagger_\mathrm{i}\right)
  +
  \left(2\pi - \bi{D}_\mathrm{e}\right)\bi{S}_\mathrm{i}
  \right]\sigma =
  -\left[\left(2\pi-\bi{D}_\mathrm{e}\right)
  -\bi{S}_\mathrm{e}\bi{S}_\mathrm{i}^{-1}
  \left(2\pi-\bi{D}_\mathrm{i}\right)
  \right]\esp,
  \tag{\ref{eq:full-IEF} from Chapter \ref{ch:CSM}}
\end{equation}

@@

In this Section, I will show how one can reformulate the transmission problem
in a variational fashion.
Such a formulation allows for a larger vector space, with weaker regularity
conditions, to be explored as solution space for the problem.
We will follow the exposition of \citeauthor{Ern2004-oo} quite closely
in introducing the \emph{weak} formulation of \glspl{PDE}.
For simplicity, we assume Dirichlet boundary conditions for a conductor,
\ie the basic assumption behind \acrshort{COSMO}.
The \emph{strong} formulation of the electrostatic problem then reads:
\begin{equation}\label{eq:poisson-cosmo}
  \nabla^2\esp = -4\pi\rho,
  \quad \esp \in \mathcal{C}^2_0(\Omegai),
\end{equation}
where $\mathcal{C}^2_0(\Omegai)$ is the vector space of twice
continuously differentiable functions in $\Omegai$ with null trace on
$\Gamma$.
We can relax this regularity requirement on $\Psi$ by introducing the
\emph{Hilbert Sobolev space} of test functions $H_0^1(\Omegai)$:
\begin{equation}
  H_0^1(\Omegai) = \lbrace
  f : \Omegai \rightarrow \mathbb{R}
  | f, \nabla f \in L^2(\Omegai), f_\Gamma = 0
  \rbrace.
\end{equation}
Projecting the differential problem onto this space and using the
$\eta\nabla^2\gamma = \nabla\cdot(\eta\nabla\gamma) - \nabla\eta\cdot\nabla\gamma$ identity
one obtains the \emph{weak} formulation of the differential problem:
\begin{equation}\label{eq:poisson-weak}
  \left\{
  \begin{aligned}
    &\text{Seek $\Psi \in H_0^1(\Omegai)$ such that:}\\
    &(\nabla\zeta, \nabla\Psi) =
    -4\pi(\zeta, \rho) \quad
    \forall \zeta\in H_0^1(\Omegai)
  \end{aligned}
  \right.
\end{equation}
The form $(\nabla\cdot, \nabla\cdot)$ is \emph{bilinear} and continuous
in $H_0^1(\Omegai)\times H_0^1(\Omegai)$, while $-4\pi(\cdot, \rho)$ is
\emph{linear} and continuous in $H_0^1(\Omegai)$.
We rewrite Eq. \eqref{eq:poisson-weak} in the abstract form:
\begin{equation}
  \left\{
  \begin{aligned}
    &\text{Seek $u \in V$ such that:}\\
    &\forall v \in V \,\,\,\,\,
    a(u, v) = b(v)
  \end{aligned}
  \right.
  \tag{\ref{eq:weak}}
\end{equation}
and state the following fundamental results:
\begin{lemma}[Lax--Milgram]
  If the bilinear form $a$ is \emph{continous} and \emph{coercive}
  in $V$, then, for any continuous linear form $b$, Problem~\eqref{eq:weak} is
  \emph{well-posed}.
\end{lemma}
\begin{corollary}[Variational property]
  If the bilinear form is \emph{symmetric} and \emph{positive}
  the unique solution to Problem \eqref{eq:weak} is
  the \emph{unique minimum} on $V$ of the functional:
  \[
  \mathcal{F}(u) = \frac{1}{2}a(u, u) - b(u)
  \]
\end{corollary}

@@

\todo[inline]{Write this section}
Approximability, conformity, consistency and asymptotic consistency
of the approximation setting, $\inf$-$\sup$ condition, Céa's Lemma
and generalized $\inf$-$\sup$ condition.\autocite{Ern2004-oo}

@@

\begin{equation}
  \begin{pmatrix}
   \sigma^\mathrm{tot} \\
   \kappa^\mathrm{tot}
  \end{pmatrix}
  =
  \begin{pmatrix}
   \sigma_\mathrm{HF} \\
   \kappa_\mathrm{HF}
  \end{pmatrix}
  +
  \begin{pmatrix}
   \sigma \\
   \kappa
  \end{pmatrix}
\end{equation}
and similarly for the sources:
\begin{equation}
\begin{aligned}
  \begin{pmatrix}
   \esp(\tamp{}, \tbar{}) \\
   \zeta(\tamp{}, \tbar{})
  \end{pmatrix}
  &=
  \begin{pmatrix}
   \esp_\mathrm{HF} \\
   \zeta_\mathrm{HF}
  \end{pmatrix}
  +
  \begin{pmatrix}
    \esp_{\mathrm{N}}(\tamp{}, \tbar{}) \\
    \zeta_{\mathrm{N}}(\tamp{}, \tbar{})
  \end{pmatrix} \\
  &=
  \begin{pmatrix}
  \braket{\mathrm{HF} | \esp  | \mathrm{HF}} \\
  \braket{\mathrm{HF} | \zeta | \mathrm{HF}}
  \end{pmatrix}
  +
  \begin{pmatrix}
   \braket{\mathrm{HF} | \tilde{\esp} | \mathrm{HF}}
  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{\esp} | \mathrm{HF}} \\
   \braket{\mathrm{HF} | \tilde{\zeta} | \mathrm{HF}}
  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{\zeta} | \mathrm{HF}}
  \end{pmatrix}
\end{aligned}
\end{equation}
The \acrshort{BCH} expansions of the source terms have been
rewritten as:
\begin{subequations}
  \begin{align}
  \overline{\esp} &= \esp + \tilde{\esp} = \esp + \commutator{\esp}{T}
  + \frac{1}{2}\commutator{\commutator{\esp}{T}}{T} \\
  \overline{\kappa} &= \kappa + \tilde{\kappa} = \kappa + \commutator{\kappa}{T}
  + \frac{1}{2}\commutator{\commutator{\kappa}{T}}{T}
  \end{align}
\end{subequations}
taking advantage of the fact that both $\esp$ and $\kappa$ are
nondiagonal, one-electron operators and hence their commutator
expansions truncate at the third term. Moreover, normal-ordering of the
operators has been introduced.\autocite{Crawford2000-ey, Shavitt2009-mr}
This is equivalent to applying a shift to the correlation part of the
source term one-electron operators to remove the reference source terms:
\begin{equation}
  \begin{aligned}
  \begin{pmatrix}
    \esp_{\mathrm{N}}(\tamp{}, \tbar{}) \\
    \zeta_{\mathrm{N}}(\tamp{}, \tbar{})
  \end{pmatrix}
  &=
  \begin{pmatrix}
   \braket{\mathrm{HF} | \tilde{\esp} | \mathrm{HF}}
  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{\esp} | \mathrm{HF}} \\
   \braket{\mathrm{HF} | \tilde{\zeta} | \mathrm{HF}}
  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{\zeta} | \mathrm{HF}}
  \end{pmatrix}
  \\
  &=
  \begin{pmatrix}
    \braket{\mathrm{HF} | \overline{\esp} - \esp_\mathrm{HF} | \mathrm{HF}}
  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{\esp} - \esp_\mathrm{HF} | \mathrm{HF}} \\
  \braket{\mathrm{HF} | \overline{\zeta} - \zeta_\mathrm{HF} | \mathrm{HF}}
  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{\zeta} -
  \zeta_\mathrm{HF}| \mathrm{HF}}
  \end{pmatrix}
\end{aligned}
\end{equation}
